Global Climate and Atmospheric Composition of the Ultra-Hot Jupiter WASP-103\lowercase{b} from \HST\ and \Spitzer\ Phase Curve Observations

### 2018-05-04 arxiv_v2/ms.tex
arxiv v2

@@

\subsection{Photometric Monitoring}
To assess how stellar activity might impact the phase curve observations, we monitored WASP-103's photometric variability over 158 nights during 2014 - 2016 with the Tennessee State University Celestron 14-inch (C14) automated imaging telescope (AIT), located at Fairborn Observatory in southern Arizona \citep[][]{henry99}.  The observations of WASP-103 were made in the Cousins R passband with an SBIG STL-1001E CCD camera.  Each observation consisted of 4--10 consecutive exposures on WASP-103 along with several dozen comparison stars in the same field. The individual consecutive frames were co-added and reduced to differential magnitudes (i.e., WASP-103 minus the mean brightness of the six best comparison stars). The nightly observations were corrected for bias, flat-fielding, and differential atmospheric extinction.  \added{We determined extinction corrections from linear, least-squares fits to seasonal plots of nightly differential magnitudes versus air mass and computing the slope in magnitudes per airmass.}

@@

To maximize the possibility of detecting WASP-103's rotation, we normalized the
photometry such that each observing season has the same mean, thereby removing
any long-term variability in WASP-103 and/or the comparison stars
(Figure\,\ref{fig:photometry}, top panel).  To estimate the stellar rotation
period, we performed a periodogram analysis of the normalized data set based on
least-squares fitting of sine curves.  The resulting frequency spectrum and the
phase curve computed with the best period are shown in the middle and lower
panels of Figure\,\ref{fig:photometry} respectively. The best-fit period is
6.814 days, which agrees closely with the estimated stellar rotation period of
6.855 days \citep[based on the projected stellar rotation velocity and stellar
radius reported in][]{gillon14}. \added{There are two nearby peaks in the
periodogram (panel b of Figure\,\ref{fig:photometry}) that are one-year aliases
of each other, and we chose the peak that better matches the stellar rotation velocity.} The peak-to-peak variability amplitude is 0.005 mag.  Based on the formalism in \cite{zellem17}, we calculate that this variability will bias the measured eclipse depth by $\lesssim10$\,parts per million (ppm) from epoch to epoch, which is well below the photon-limited precision of our measurements.

@@

\begin{figure*}
\includegraphics[width = 1.0\textwidth]{fig5.pdf}
\caption{WASP-103b phase curve observations from \HST/WFC3 (top) and
\Spitzer/IRAC (middle and bottom). For clarity, the data are phase-folded on the
planet's orbital period and binned in 30 uniformly spaced bins between 0 and 1
\added{(corresponding to 0.8 hours)}. The left column shows the phase curves with systematic noise removed (black points) compared to the best fit spherical harmonics model (colored lines). The error bars denote $1\,\sigma$ uncertainties (in some cases, the errors are smaller than the data points).  We include the transits in the fit, but they are not displayed in this figure. The right-hand column shows the binned residuals for the best-fit light curve. The gray error bars in the upper right of the left panels correspond to 500 ppm, to illustrate the changing y-axis scale.}
\label{fig:phasecurves}
\end{figure*}

@@

\subsection{Comparison with Previous Eclipse Observations}
We compared our results to the dayside emission spectrum reported by \cite{cartier17}, which is based on a Gaussian process analysis of two secondary eclipses from \HST/WFC3. The shape of their spectrum is consistent with what we find, but their eclipse depths are 125 ppm smaller on average (a difference of about 10\%). A likely explanation for this difference is that the \cite{cartier17} analysis does not include the planet's thermal phase variation, so that the Gaussian process models it as an instrument systematic.  If the phase variation is absorbed into the systematic model, the measured eclipse depths would be biased low.  \added{By visual inspection} of Figure\,\ref{fig:model_comparison}, we estimate the amplitude of this effect is $\sim100$ ppm, which is comparable to the offset between the two analyses. Our estimated uncertainties are a factor of four smaller than those reported in \cite{cartier17}, which is consistent with photon-limited expectations (our data set includes two additional eclipses, a factor of five longer out-of-eclipse baseline, and has 60\% larger wavelength bins).

@@

\added{As a test, we also calculated the difference in brightness temperature
between the \HST\ and \Spitzer\ 4.5 $\mu$m data (reported in
Table\,\ref{table:teffs}) for both nightside phases ($\phi = 0.1$ and 0.9). At
phase 0.1 and 0.9, the \Spitzer\ temperature is lower by 2.7 and 5.6\,$\sigma$,
respectively.  These values are a lower limit to the significance, because the
brightness temperatures noted in the table also include the uncertainty in the
stellar $T_\mathrm{eff}$ (which increases the uncertainty on the absolute planet
temperature but not the relative temperatures that are relevant for this
calculation). The drop in brightness temperature is more significant at phase 0.9 than it is at 0.1, providing further evidence in addition to the phase 0.1 retrieval that the nightside temperature pressure profile is not inverted.}

@@

\begin{figure*}
\includegraphics[width = 1.0\textwidth]{fig17.pdf}
\caption{Flux-calibrated spectra for WASP-103b (left column) compared to
observed spectra for brown dwarfs (middle) and directly imaged companions
(right), assuming a distance of 10 pc. The WASP-103b spectra are from the
nightside (phase 0.1, top row), quadrature (phase 0.25, middle) and the dayside
(phase 0.5, bottom). Each row shows spectra from objects of comparable
temperature. The dotted gray line corresponds to the best fit blackbody.
Effective temperatures are listed in the upper right corners. The wavelength
bins used to calculate the water feature amplitude $A_1$ are shaded in gray.
\added{Note that the panels do not have the same y-scale, since the objects have
different radii which leads to large variation in absolute flux.}}
\label{fig:planetstarcomparison}
\end{figure*}

>>>>

\author{Thomas G.\ Beatty}
\affiliation{Department of Astronomy \& Astrophysics, The Pennsylvania State University, 525 Davey Lab, University Park, PA 16802}
\affiliation{Center for Exoplanets and Habitable Worlds, The Pennsylvania State University, 525 Davey Lab, University Park, PA 16802}

@@

\subsection{Photometric Monitoring}
To assess how stellar activity might impact the phase curve observations, we monitored WASP-103's photometric variability over 158 nights during 2014 - 2016 with the Tennessee State University Celestron 14-inch (C14) automated imaging telescope (AIT), located at Fairborn Observatory in southern Arizona \citep[][]{henry99}.  The observations of WASP-103 were made in the Cousins R passband with an SBIG STL-1001E CCD camera.  Each observation consisted of 4--10 consecutive exposures on WASP-103 along with several dozen comparison stars in the same field. The individual consecutive frames were co-added and reduced to differential magnitudes (i.e., WASP-103 minus the mean brightness of the six best comparison stars). The nightly observations were corrected for bias, flat-fielding, and differential atmospheric extinction.  We determined extinction corrections from linear, least-squares fits to seasonal plots of nightly differential magnitudes versus air mass and computing the slope in magnitudes per airmass.

@@

To maximize the possibility of detecting WASP-103's rotation, we normalized the
photometry such that each observing season has the same mean, thereby removing
any long-term variability in WASP-103 and/or the comparison stars
(Figure\,\ref{fig:photometry}, top panel).  To estimate the stellar rotation
period, we performed a periodogram analysis of the normalized data set based on
least-squares fitting of sine curves.  The resulting frequency spectrum and the
phase curve computed with the best period are shown in the middle and lower
panels of Figure\,\ref{fig:photometry} respectively. The best-fit period is
6.814 days, which agrees closely with the estimated stellar rotation period of
6.855 days \citep[based on the projected stellar rotation velocity and stellar
radius reported in][]{gillon14}. There are two nearby peaks in the
periodogram (panel b of Figure\,\ref{fig:photometry}) that are one-year aliases
of each other, and we chose the peak that better matches the stellar rotation velocity. The peak-to-peak variability amplitude is 0.005 mag.  Based on the formalism in \cite{zellem17}, we calculate that this variability will bias the measured eclipse depth by $\lesssim10$\,parts per million (ppm) from epoch to epoch, which is well below the photon-limited precision of our measurements.

@@

\begin{figure*}
\includegraphics[width = 1.0\textwidth]{fig5.pdf}
\caption{WASP-103b phase curve observations from \HST/WFC3 (top) and
\Spitzer/IRAC (middle and bottom). For clarity, the data are phase-folded on the
planet's orbital period and binned in 30 uniformly spaced bins between 0 and 1
(corresponding to 0.8 hours). The left column shows the phase curves with systematic noise removed (black points) compared to the best fit spherical harmonics model (colored lines). The error bars denote $1\,\sigma$ uncertainties (in some cases, the errors are smaller than the data points).  We include the transits in the fit, but they are not displayed in this figure. The right-hand column shows the binned residuals for the best-fit light curve. The gray error bars in the upper right of the left panels correspond to 500 ppm, to illustrate the changing y-axis scale.}
\label{fig:phasecurves}
\end{figure*}

@@

\subsection{Comparison with Previous Eclipse Observations}
We compared our results to the dayside emission spectrum reported by \cite{cartier17}, which is based on a Gaussian process analysis of two secondary eclipses from \HST/WFC3. The shape of their spectrum is consistent with what we find, but their eclipse depths are 125 ppm smaller on average (a difference of about 10\%). A likely explanation for this difference is that the \cite{cartier17} analysis does not include the planet's thermal phase variation, so that the Gaussian process models it as an instrument systematic.  If the phase variation is absorbed into the systematic model, the measured eclipse depths would be biased low.  By visual inspection of Figure\,\ref{fig:model_comparison}, we estimate the amplitude of this effect is $\sim100$ ppm, which is comparable to the offset between the two analyses. Our estimated uncertainties are a factor of four smaller than those reported in \cite{cartier17}, which is consistent with photon-limited expectations (our data set includes two additional eclipses, a factor of five longer out-of-eclipse baseline, and has 60\% larger wavelength bins).

@@

As a test, we also calculated the difference in brightness temperature
between the \HST\ and \Spitzer\ 4.5 $\mu$m data (reported in
Table\,\ref{table:teffs}) for both nightside phases ($\phi = 0.1$ and 0.9). At
phase 0.1 and 0.9, the \Spitzer\ temperature is lower by 2.7 and 5.6\,$\sigma$,
respectively.  These values are a lower limit to the significance, because the
brightness temperatures noted in the table also include the uncertainty in the
stellar $T_\mathrm{eff}$ (which increases the uncertainty on the absolute planet
temperature but not the relative temperatures that are relevant for this
calculation). The drop in brightness temperature is more significant at phase 0.9 than it is at 0.1, providing further evidence in addition to the phase 0.1 retrieval that the nightside temperature pressure profile is not inverted.

@@

\begin{figure*}
\includegraphics[width = 1.0\textwidth]{fig17.pdf}
\caption{Flux-calibrated spectra for WASP-103b (left column) compared to
observed spectra for brown dwarfs (middle) and directly imaged companions
(right), assuming a distance of 10 pc. The WASP-103b spectra are from the
nightside (phase 0.1, top row), quadrature (phase 0.25, middle) and the dayside
(phase 0.5, bottom). Each row shows spectra from objects of comparable
temperature. The dotted gray line corresponds to the best fit blackbody.
Effective temperatures are listed in the upper right corners. The wavelength
bins used to calculate the water feature amplitude $A_1$ are shaded in gray.
Note that the panels do not have the same y-scale, since the objects have
different radii which leads to large variation in absolute flux.}
\label{fig:planetstarcomparison}
\end{figure*}
